### 2023-04-27 paper.Rmd
editing

  write.table(
    df, path, row.names = FALSE, sep = ","
  )
} else {
  df <- read.csv(path)
}

@@

>>>>

  write.table(
    df, path, row.names = FALSE, sep = ","
  )

  # investigate the best 0.5% of k-means anticlustering partitions, for discussion
  # best_0.1percent_anticlusterings <- df$kmeans /max(df$kmeans) > .999
  # best_partitions <- partitions[best_0.1percent_anticlusterings]
  # 
  # neighbor_matrix <- anticlust:::selection_matrix_from_clusters
  # neighbourhoods <- lapply(best_partitions, neighbor_matrix)
  # neighbourhoods <- lapply(neighbourhoods, function(x) as.integer(x[lower.tri(x)]))
  # pairs <- Reduce("+", neighbourhoods)
  # length(pairs)
  # length(neighbourhoods[[1]])
  # choose(14, 2)
  # min(pairs) # 40
  # max(pairs) # 95
  # length(best_partitions) # 146
  # -> all items are grouped with one another, and no items are always grouped with one another

} else {
  df <- read.csv(path)
}

@@

## Closing remarks on the nature of anticlustering

```{r}
df <- read.csv("partition_objectives.csv")
```

Traditionally, anticlustering is implemented by reversing an objective function used in classical cluster analysis. Whereas cluster analysis seeks homogeneity within groups and separation between groups, the reversal results in heterogeneity within groups and similarity between groups. While tightly linked in a mathematical sense, an important semantic difference is that cluster analysis is concerned with finding "true" structure in the data, e.g., by defining a taxonomy on naturally occurring groups [@everitt1979]. Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern. Correspondingly, different similarly adequate anticlustering partitions tend to look very different, while in cluster analysis, similar elements always tend to be grouped in the same cluster. In Figure 1, for example, we observe a strong concentration of anticlustering partitions: An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution. Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not; the minimum amount of co-occurrence is 40 and the maximum amount of co-occurrence is 95. Thus, a high anticlustering objective does not depend on assigning individual items either to the same or a different group. In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%. The clustering quality crucially depends on assigning similar elements to the same group. Hence, while there are many different high quality anticlustering partitions, there are only very few adequate clustering partitions, constituting another relevant difference between cluster analysis and anticlustering. Interestingly, the k-plus objective further loosens the link between cluster analysis and anticlustering. Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering in the usual sense. Thus, even though k-plus extends the classical k-means clustering criterion, it is a pure anticlustering objective.

An interesting mixture of cluster analysis and anticlustering is fair clustering [e.g. @chhabra2021; @delBarrio2022]. Fair clustering performs a cluster analysis under the restriction that the distribution of a set of "protected" attributes---such as race, sex or age---should be similar between clusters. Thus, as compared to pure cluster analysis, fair clustering biases the identification of groups with regard to the protected attributes, striving for algorithmic fairness. However, a relevant difference to anticlustering is that fair clustering is still concerned with identifying "true" structure in the data [@delBarrio2022], while anticlustering is not concerned with identifying the underlying structure of a data set. However, these methods are similar because anticlustering logically corresponds to fair clustering when only the protected attributes are considered. Hence, a fruitful outlook for research might be to investigate if approaches stemming from the literature on fair clustering can be adopted for anticlustering, and vice versa.
